\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-10-08 adaptive-estimation.tex
fixed (I think) the proof of Theorem 4

\subsection{Maximal Efficiency using One Round of Threshold Adaptation}
In Section~\ref{sec:preliminary} we considered an estimator that is based on binary messages of the form 
\[
B_i = \mathbf 1_{X_i > \theta_0},\qquad i=1,\ldots,n,
\]
and deduced that it is asymptotically normal with variance $1/\eta(\theta-\theta_0)$.
%
We now show that a similar encoding leads to an asymptotically normal estimator attaining the lower variance bound $1/\eta(0)$, provided we allow to  update the threshold value $\theta_0$ based on previously observed bits at least once. 
%
In this procedure we separate the sample into two disjoint sets: $X_1,\ldots,X_{n_1}$ and $X_{n_1+1},\ldots,X_n$ for some $n_1 < n$.
We first use the estimator \eqref{eq:estimator_naive} to obtain an estimate ${\theta}_{n_1}$ based on $B_1,\ldots,B_{n_1}$, and then use ${\theta}_{n_1}$ as the new threshold value to obtain messages $B_{n_1+1}, \ldots, B_n$. Figure~\ref{fig:one_round} illustrates a diagram of this procedure. 
%
The specific encoding and estimation scheme, as well as its asymptotic performance, are given by the following theorem:
\begin{thm}
For $i=1,\ldots,n$ set
\[
B_i = \begin{cases}
 \mathbf 1_{X_i \geq \theta_0} & i = 1,\ldots,n_1, \\
 \mathbf 1_{X_i \geq {\theta}_{n_1} }& i={n_1+1,\ldots,n},
\end{cases}
\]
where $n = n_1+ n_2$ and 
\[
{\theta}_{n_1} \triangleq \theta_0 + F^{-1}\left(
\frac{1}{n_1} \sum_{i=1}^{n_1} B_i 
 \right).
\] 
Let 
\[
{\theta}_{n_2}  = {\theta}_{n_1} +  F^{-1} \left( \frac{1}{n_2} \sum_{i=n_1+1}^{n_2} B_i \right)
\]
and assume that, as $n\to \infty$, $n_1(n) \rightarrow \infty$ and $n_2(n)/n \rightarrow 1$. Then:
\begin{align*}
 \sqrt{n} \left( {\theta}_{n_2} - \theta  \right) \overset{d}{\longrightarrow}  \Ncal\left( 0, 1/\eta(0) \right).
\end{align*}
\end{thm}
%
%
\begin{proof}
By the law of large numbers,
\[
p_{n_1} \triangleq \frac{1}{n_1} \sum_{i=1}^{n_1} B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),
\]
so that ${\theta}_{n_1}$ converges almost surely to $\theta$ as $n$ goes to infinity. 
For $t\in \mathbb R$, set
\[
p_{n_2}(t) \triangleq \frac{1}{n_2} \sum_{i=n_1+1}^{n} \mathbf 1_{X_i \geq t}. 
\]
Since $n_2(n)/n \to 1$, the central limit theorem and Slutsky's theorem imply
\begin{align*}
\sqrt{n} \left( p_{n_2}(t) - F( \theta - t) \right)   \overset{d}{\rightarrow} \sqrt{F(\theta - t) F \left( t - \theta\right)} Z,
\end{align*}
where $Z \sim \Ncal(0,1)$. In addition, Slutsky's theorem imply
\[
\sqrt{F(\theta-\theta_{n_1}) F(\theta_{n_1} - \theta)} Z_1 \overset{d}{\rightarrow}  F(0) Z,  
\]
hence 
\[
\sqrt{n} \left( p_{n_2}( \theta_{n_1}) - F( \theta - \theta_{n_1}) \right)  \overset{d}{\rightarrow} F(0) Z.
\]
Finally, we use the delta method with $g(x) = F^{-1}(x)$ to conclude
\begin{align*}
& \sqrt{n}\left( \theta_{n_2} - \theta \right) = \sqrt{n} \left( \theta_{n_1} + F^{-1}(p_{n_2}(\theta_{n_1})) - \theta \right) \\
&= \sqrt{n} \left( g(p_{n_2}(\theta_{n_1})) - g \left(F(\theta-\theta_{n_1}) \right)  \right) \overset{d}{\to} \Ncal\left(0, 
F^2(0)/ f^2(0) \right).
\end{align*}
\end{proof}
%

>>>>

\subsection{Maximal Efficiency using One Round of Threshold Adaptation}
In Section~\ref{sec:preliminary} we considered an estimator that is based on binary messages of the form 
\[
B_i = \mathbf 1_{X_i > \theta_0},\qquad i=1,\ldots,n,
\]
and deduced that it is asymptotically normal with variance $1/\eta(\theta-\theta_0)$.
%
We now show that a similar encoding leads to an asymptotically normal estimator attaining the lower variance bound $1/\eta(0)$, provided we allow to  update the threshold value $\theta_0$ based on previously observed bits at least once. 
%
In this procedure we separate the sample into two disjoint sets: $X_1,\ldots,X_{n_1}$ and $X_{n_1+1},\ldots,X_n$ for some $n_1 < n$.
We first use the estimator \eqref{eq:estimator_naive} to obtain an estimate ${\theta}_{n_1}$ based on $B_1,\ldots,B_{n_1}$, and then use ${\theta}_{n_1}$ as the new threshold value to obtain messages $B_{n_1+1}, \ldots, B_n$. Figure~\ref{fig:one_round} illustrates a diagram of this procedure. 
%
The specific encoding and estimation scheme, as well as its asymptotic performance, are given by the following theorem:
\begin{thm}
For $i=1,\ldots,n$ set
\[
B_i = \begin{cases}
 \mathbf 1_{X_i \geq \theta_0} & i = 1,\ldots,n_1, \\
 \mathbf 1_{X_i \geq {\theta}_{n_1} }& i={n_1+1,\ldots,n},
\end{cases}
\]
where $n = n_1+ n_2$ and 
\[
{\theta}_{n_1} \triangleq \theta_0 + F^{-1}\left(
\frac{1}{n_1} \sum_{i=1}^{n_1} B_i 
 \right).
\] 
Let 
\[
{\theta}_{n_2}  \triangleq {\theta}_{n_1} +  F^{-1} \left( \frac{1}{n_2} \sum_{i=n_1+1}^{n_2} B_i \right)
\]
and assume that, as $n\to \infty$, $n_1(n) \rightarrow \infty$ and $n_2(n)/n \rightarrow 1$. Then:
\begin{align*}
 \sqrt{n} \left( {\theta}_{n_2} - \theta  \right) \overset{d}{\longrightarrow}  \Ncal\left( 0, 1/\eta(0) \right).
\end{align*}
\end{thm}
%
%
\begin{proof}
For $t\in \mathbb R$, set
\[
p_{n_2}(t) \triangleq \frac{1}{n_2} \sum_{i=n_1+1}^{n} \mathbf 1_{X_i \geq t}. 
\]
Since $n_2(n)/n \to 1$, if follows from the Berry Esseen theorem that there exists $C$, independent of $n$ and $t$, such that 
\begin{align*}
\sup_{t \in \R} \left| \sqrt{n} \frac{ p_{n_2}(t) - F( \theta - t)} {F(\theta - t) F \left( t - \theta\right) }  - \Phi(t) \right|  \leq C/\sqrt{n}. 
\end{align*}
where $Z \sim \Ncal(0,1)$. It follows that
\begin{align}
\label{eq:two_step_proof}
 \sqrt{n} \frac{ p_{n_2}(Y_n) - F(0)} {F^2(0) }  \overset{d}{\rightarrow} \Ncal(0,1)
\end{align}
for any sequence of random variables $\{Y_n\}$ converging in probability to $\theta$. In particular, since
\[
p_{n_1} \triangleq \frac{1}{n_1} \sum_{i=1}^{n_1} B_i \overset{a.s.}{\rightarrow} F(\theta - \theta_0),
\]
by the law of large numbers, \eqref{eq:two_step_proof} holds with $Y_n = \theta_{n_1}$. Finally, we use the delta method with $g(x) = F^{-1}(x)$ to conclude
\begin{align*}
& \sqrt{n}\left( \theta_{n_2} - \theta \right) = \sqrt{n} \left( \theta_{n_1} + F^{-1}(p_{n_2}(\theta_{n_1})) - \theta \right) \\
&= \sqrt{n} \left( g(p_{n_2}(\theta_{n_1})) - g \left(F(\theta-\theta_{n_1}) \right)  \right) \overset{d}{\to} \Ncal\left(0, 
F^2(0)/ f^2(0) \right).
\end{align*}
\end{proof}
%
